Persistence in functional topology: local-connectivity, q-tameness and a correction to a theorem of Morse

### 2021-04-15 introduction.tex
Improve abstract

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Now a days, when thinking about the pioneering work of Marston Morse in this domain, a smooth function on a closed differential manifold is possibly our first association but, as we describe next, it should not be our last.
In the \textit{finite} context, Morse theory was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}; where he also move beyond this setting to give a new proof of Bott's periodicity, applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another example of the importance of Morse's insights in infinite contexts is Floer's work on the Arnold conjecture.
Morse himself worked in a significantly less constraint setup, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context of semi-continuous functionals on metric spaces.
He name this body of work \textit{functional topology}.
A motivating problem for him came from the study of minimal surfaces.
In 1931, Jesse Douglas solved Plateau’s Problem \cite{Douglas.1931} for which he received the fields medal in 1936.
The problem amounts to finding for a given admissible boundary curve $\Gamma$ an area-minimizing embedded surface $\Sigma$ with $\partial \Sigma = \Gamma$.
In \cite{Morse.1939, Morse.1941}, Morse and Tompkins used functional topology to give conditions for the existence of unstable minimal surfaces, i.e., critical points of the area functional that do not locally minimize area.

>>>>

The interplay between the critical set of a function and the topology of its domain is a cornerstone of modern mathematics.
Now a days, when thinking about the pioneering work of Marston Morse in this domain, a smooth function on a closed differential manifold is possibly our first association but, as we describe next, it should not be our last.
In the \textit{finite} context, Morse theory was masterfully presented in Milnor's famous book on the subject \cite{MR0163331}; where he also move beyond this setting to give a new proof of Bott's periodicity, applying Morse theory to the energy functional of paths in a Riemannian manifold.
Another example of the importance of Morse's insights in infinite contexts is Floer's work on the Arnold conjecture.
Morse himself worked in a significantly less constraint setup, publishing in the 1930s a series of papers \cite{Morse.1937, Morse.1938, Morse.1940, MR9102} in which he established the key results of Morse theory in the broad context of semi-continuous functionals on metric spaces.
He name this body of work \textit{functional topology}.
Motivation for its development came from the study of minimal surfaces.
Jesse Douglas received the fields medal in 1936 for solving in 1931 \cite{Douglas.1931} Plateau’s Problem: prove the existence of a bounding area-minimizing surface for any given admissible curve.
In \cite{Morse.1939, Morse.1941}, Morse and Tompkins used functional topology to give conditions for the existence of \textit{unstable} minimal surfaces, i.e., critical points of the Douglas functional that do not locally minimize area.
